From DSL to HPC Component-Based Runtime:\\ A Multi-Stencil DSL Case Study

### 2015-07-31 WOLFHPC15/article/concept.tex

%----------------------------------------
\subsection{A Multi-Stencil Formalization}
\label{sect:multistencil}
%----------------------------------------
Let introduce some formal definitions used in the rest of the paper.
For $\Delta$ the set of data of the simulation, a stencil kernel $s$ is defined as the quintuplet
\begin{equation} 
s(R,w,exp,d),
\label{eq:st}
\end{equation}
where $R$ is a set of pair $(r,n)$, with $r \in \Delta$ is a data read by the computation, and $n$ is a stencil shape (a neighborhood) applied on $r$. The data written by the kernel is denoted $w \in \Delta$, $exp$ is the numerical expression of the stencil kernel. Finally the numerical expression is applied on the computation domain $d$.

For example, in Figure~\ref{fig:ex1}, assuming the computation domain (full lines) is $dc1$ and the stencil shape is $n1$, the stencil kernel can be defined with:
\begin{equation*}
R: {(B,n1)} \quad w: A \quad d: dc1
\end{equation*}
\begin{equation*}
exp: A(x,y)=B(x+1,y)+B(x-1,y)+B(x,y+1)+B(x,y-1).
\end{equation*}
On the other hand, in the example of Figure~\ref{fig:ex2}, assuming the computation domain is $dc2$ and the stencil shape is $n2$, the stencil kernel can be defined with:
\begin{equation*}
R: {(C,n2),(A,null)} \quad w: A \quad d: dc2
\end{equation*}
\begin{equation*}
exp: A(x,y)=A(x,y)+C(x1,y1)+C(x1+1,y1)
\end{equation*}
One can notice that the input data $A$ is associated to the stencil shape $null$. This means that no stencil shape is applied on $A$, but only local accesses.

A local, or auxiliary, computation is a particular case of stencil kernel without neighborhood information in the numerical expression, which makes this kind of computation purely local. A local computation $l$ is defined as the quadruplet:
\begin{equation} 
l(R_l,w,exp,d),
\label{eq:loc}
\end{equation}
where $R_l$ is the set of input data, and other parameters are the same than for a stencil computation.

>>>>

%----------------------------------------
\subsection{A Multi-Stencil Formalization}
\label{sect:multistencil}
%----------------------------------------
This section introduces some formal definitions used in the rest of the paper.
Let $\Delta$ be the set of data of the simulation
A stencil kernel $s$ is defined as the quintuplet
\begin{equation} 
s(R,w,exp,d),
\label{eq:st}
\end{equation}
where $R$ is a set of pairs $(r,n)$, with $r \in \Delta$ a data read by the computation, and $n$ the stencil shape (neighborhood) used to access $r$. The data written by the kernel is denoted $w \in \Delta$. $exp$ is the numerical expression of the stencil kernel. Finally the numerical expression is applied on the computation domain $d$.

For example, in Figure~\ref{fig:ex1}, assuming the computation domain (full lines) is $dc1$ and the stencil shape is $n1$, the stencil kernel can be defined as:
\begin{equation*}
R: \{(B,n1)\} \quad w: A \quad d: dc1
\end{equation*}
\begin{equation*}
exp: A(x,y)=B(x+1,y)+B(x-1,y)+B(x,y+1)+B(x,y-1).
\end{equation*}
On the other hand, in the example of Figure~\ref{fig:ex2}, assuming the computation domain is $dc2$ and the stencil shape is $n2$, the stencil kernel can be defined as:
\begin{equation*}
R: \{(C,n2),(A,local)\} \quad w: A \quad d: dc2
\end{equation*}
\begin{equation*}
exp: A(x,y)=A(x,y)+C(x1,y1)+C(x1+1,y1)
\end{equation*}
One can notice that the input data $A$ is associated to the stencil shape $local$. This means that the stencil shape applied on $A$ is limited to the element with the exact same coordinate as the result element.

The specificities of kernels whose stencil shapes are all local makes it possible to apply specific optimizations.
We therefore provide a specific definition in this case.
A local (or auxiliary) kernel $l$ is defined as the quadruplet:
\begin{equation} 
l(R_l,w,exp,d),
\label{eq:loc}
\end{equation}
where $R_l$ is the set of input data, and other parameters are the same as for normal stencil kernels.
